---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# RSQLServer

[![CRAN](http://www.r-pkg.org/badges/version/RSQLServer)](http://cran.r-project.org/package=RSQLServer)
[![Travis-CI build status](https://travis-ci.org/imanuelcostigan/RSQLServer.svg?branch=master)](https://travis-ci.org/imanuelcostigan/RSQLServer)
[![Appveyor build status](https://ci.appveyor.com/api/projects/status/i2oedybuqi2o5crg/branch/master?svg=true)](https://ci.appveyor.com/project/imanuelcostigan/rsqlserver/branch/master)
[![Coverage status](https://codecov.io/gh/imanuelcostigan/RSQLServer/branch/master/graph/badge.svg)](https://codecov.io/gh/imanuelcostigan/RSQLServer)

An R package that provides a SQL Server R Database Interface ([DBI](https://github.com/rstats-db/DBI)), based on the cross-platform [jTDS JDBC driver](http://jtds.sourceforge.net/index.html).

## Installation

You can't install the package from CRAN yet. But you install the development version from GitHub:

```{r, eval=FALSE}
# install.packages('devtools')
devtools::install_github('imanuelcostigan/RSQLServer')
```

## Config file

We recommend that you store server details and credentials in `~/sql.yaml`. This is partly so that you do not need to specify a username and password in calls to `dbConnect()`. But it is also because in testing, we've found that the jTDS single sign-on (SSO) library is a bit flaky. The contents of this file should look something like this:

```yaml
SQL_PROD:
    server: 11.1.111.11
    type: &type sqlserver
    port: &port 1433
    domain: &domain companyname
    user: &user winusername
    password: &pass winpassword
    useNTLMv2: &ntlm true
SQL_DEV:
    server: 11.1.111.15
    type: *type
    port: *port
    domain: *domain
    user: *user
    password: *pass
    useNTLMv2: *ntlm
```

## Usage

Ensure that your `~/sql.yaml` file contains a valid SQL Server entry named `TEST`. In the following, the `TEST` server, generously provided by Microsoft for the purposes of this package's development, has a database containing the `nycflights13` package data sets.

### DBI usage

The following illustrates how you can make use of the DBI interface. Note that we **do not** attach the `RSQLServer` package.

```{r}
library(DBI)
nycflights <- dbConnect(RSQLServer::SQLServer(), server = "TEST", database = 'DBItest')
# RSQLServer only returns tables with type TABLE and VIEW.
dbListTables(nycflights)
dbListFields(nycflights, 'airlines')
dbReadTable(nycflights, 'airlines')

# Fetch all results
res <- dbSendQuery(nycflights, 'SELECT TOP 10 * FROM airlines')
dbFetch(res)
dbClearResult(res)

```

### dplyr usage

The following illustrates how you can make use of the dplyr interface. Again, we **do not** attach the `RSQLServer` package.

```{r}
library(dplyr, warn.conflicts = FALSE)
flights <- tbl(nycflights, "flights")
flights %>% 
  filter(carrier == "UA") %>% 
  arrange(-year, -month, -day, -dep_time)
collect(flights)
```

Then close the connection

```{r}
dbDisconnect(nycflights)
```

